Representation theory of Finite Groups

### 2016-11-16 Lecture note - Group Theory in Physics/serge lang/s-lang.tex
Split Lang's representation theory away from lecture notes. Noting about section 18.1 of Lang's representation of finite groups.

@@

>>>>

% The entire content of this work (including the source code
% for TeX files and the generated PDF documents) by 
% Hongxiang Chen (nicknamed we.taper, or just Taper) is
% licensed under a 
% Creative Commons Attribution-NonCommercial-ShareAlike 4.0 
% International License (Link to the complete license text:
% http://creativecommons.org/licenses/by-nc-sa/4.0/).
\documentclass{article}

% My own physics package
% The following line load the package xparse with additional option to
% prevent the annoying warnings, which are caused by the package
% "physics" loaded in package "physicist-taper".
\usepackage[log-declarations=false]{xparse}
\usepackage{physicist-taper}

\makenomenclature

% Show the keys for labels, replace options with "final" when done
% with editing.
\usepackage[draft,notref]{showkeys}

\title{Representation theory of Finite Groups}
\date{\today}
\author{Taper}

\begin{document}

\maketitle
\abstract{
A note for correspoding chapter in S. Lang's book.
}
\tableofcontents

@@

\begin{defi}[Sum $\rho\oplus \rho'$]
\nomenclature{Sum $\rho\oplus \rho'$}{\nomrefpage.}
If $\rho: G\to \Aut_R(E)$, $\rho':G\to \Aut_R(E')$ are two
representation of $G$, we define their sum $\rho\oplus \rho'$ to be
the representation on the direct sum $E\oplus E'$, with $\sigma\in G$
acting componentwise.
\end{defi}

Give $G$, observe the $G$-isomorphism classes of representations have
an additive monoid structure under the above direct sum, and also have
an associative multiplicative structure under the tensor product. With
the notation of representations, we denote this product by
\nomen{$\rho\otimes \rho'$}. This product is distributive with respect
to the addition (i.e. direct sum), by their definition.

\begin{defi}[Trace $\Tr_G$]
\nomenclature{Trace $\Tr_G$}{\nomrefpage.}
    If $G$ is a finite group, and $E$ is a $G$-module, then we can
    define the trace $\Tr_G:E\to E$ which is an $R$-homomorphism,
    namely
    \begin{equation}
        \Tr_G(x) = \sum_{\sigma\in G} \sigma x
    \end{equation}
\end{defi}
\begin{fact}
$\Tr_G$ belongs to $ \operatorname{inv}_G (E)$, i.e. is
fixed under the operation of any $\sigma\in G$:
\begin{align*}
    \tau \Tr_G(x) = \sum_{\sigma\in G} \tau \sigma x =
    \sum_{\sigma'\in G} \sigma' x
\end{align*}
\end{fact}
Let $E,F$ be twn $G$-modules, and $f:E\to F$ is an $R$-homomorphism of
$G$-modules, then we can easily extend this definition to define
\nomen{$\Tr_G(f)$} by
\begin{align}
    \Tr_G(f)&:E\to F \\
    \Tr_G(f) &= \sum_{\sigma\in G} [\sigma] f \nonumber
\end{align}
Also, $\Tr_G(f)$ is invariant under any $\sigma\in G$, so it is a
$G$-homomorphism, i.e.  $\Tr_G(f)\in \Hom_G(E,F)$.

\begin{prop}
    Let $G$ be a finite group and let $E,E',F,F'$ be $G$-modules, let
    \begin{equation*}
        E'\overset{\phi}{\to} E\overset{f}{\to} 
            F \overset{\psi}{\to} F'
    \end{equation*}
    be $R$-homomorphism, and asume that $\phi,\psi$ are
    $G$-homomorphisms. Then 
    \begin{equation}
        \Tr_G(\psi\circ f\circ\phi) = \psi\circ\Tr_G(f)\circ\phi
    \end{equation}
\end{prop}
\begin{proof}
    Proof is provided on the book \cite{lang-algebra}. The essential
    point is that
    \begin{equation}
        [\sigma] (\psi\circ f\circ\phi) = \sigma \psi \sigma^{-1}
        \sigma f \sigma^{-1} \sigma \phi \sigma^{-1} =
        [\sigma]\psi\circ [\sigma]f\circ [\sigma]\phi 
    \end{equation}
\end{proof}

\begin{thm}[Maschke]
    Let $G$ be a finite group of order $n$, and let $k$ be a field
    whose characteristic does not divide $n$. then the group ring
    $k[G]$ is semisimple.
\end{thm}
Recall the an object is called \nomen{semisimple} if it is the direct
sum of simple objects. A \nomen{simple} object is some ireducible
building blocks, without being too simple to be simple. \footnote{
    As remarked in nLab
    (\href{https://ncatlab.org/nlab/show/too+simple+to+be+simple}{link}),
    There is a general principle in mathematics that "A trivial object
    is too simple to be simple". For example, $1$ is not a prime
number.} A \nomen{simple ring} is a nonzero ring who has no two-sided ideals
other than the zero ideal and itself.\footnote{ I did not find the
    definition for a simple ring on Lang's book \cite{lang-algebra}.
Rather, I found it online.}

The proof of this theorem is provided in page 666, section 18.1 of
\cite{lang-algebra}. I do not understand it. So I do not put it down
here.

\section{Anchor}
\label{sec:Anchor}

\begin{thebibliography}{1}
    \bibitem{book} Zhongqi Ma, Group Theory in Physics
    \bibitem{Ludeling} Lecture Notes for physics751: Group Theory (for
    Physicists), by C Ludeling.
    \href{http://www.th.physik.uni-bonn.de/nilles/people/luedeling/grouptheory/data/grouptheorynotes.pdf}{Link}
    \bibitem{lang-algebra} Serge Lang. Algebra. Revised 3rd. Springer.
\end{thebibliography}
\printnomenclature
\section{License}
The entire content of this work (including the source code
for TeX files and the generated PDF documents) by 
Hongxiang Chen (nicknamed we.taper, or just Taper) is
licensed under a 
\href{http://creativecommons.org/licenses/by-nc-sa/4.0/}{Creative 
Commons Attribution-NonCommercial-ShareAlike 4.0 International 
License}. Permissions beyond the scope of this 
license may be available at \url{mailto:we.taper[at]gmail[dot]com}.
\end{document}
